### 2015-01-08 articles/2014/MWA-Radio-Environment/response-to-referee.txt
Implementing changes requested by referee

In the sentece: "alpha is typically -1.5 to -1.6"

Referee: "Presumably these nominal values of alpha are determined empirically?  This should be stated explicitly."

Answer: This is both empirically found as well as mathematically derived for uniform distributions of sources with a given average transmitter height. We think explaining this is a bit too much for this paper, so we have changed this sentence to "the exponent alpha is found by Offringa et al (2013b) to be typically -1.5 to -1.6."

"This excess is caused by the multi-variate nature of the data, for example because the noise level increases when the Galaxy goes..."

Referee: "Not sure "multi-variate" is the correct term here.  Perhaps non-stationarity and non-gaussanity of RFI is also a factor."

Answer:

Referee: "Such seemingly random impulsive RFI is ubiquitous - probably caused by ESD "sparking" and relay contacts."

Answer:

Figure 6.

"Are the centre channels of the PFB bands also flagged?  Otherwise, what are narrow horizontal lines?"

Answer: This is explained below the image: "[..] The horizontal flagged lines are flagged because they are 1.28-MHz subband edge or centre channels, which are unusable because of aliasing of the poly-phase filter bank and DC offsets, respectively."
These images are not correctly placed in the PASA-generated PDF, so this is a bit confusing.